From Big Data to Big Displays\\ High-Performance Visualization at Blue Brain

### 2017-05-15 paper.tex
Jupyter image
Review and ZeroBuf

\usepackage{cite}
\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{times}
%\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{amsfonts}
%\usepackage{url}
\usepackage{caption}
\usepackage{multirow}
\usepackage{paralist}
\usepackage{listings}

@@

@@

\section{Collaborative Displays}

Collaborative display systems are the integration point of the Blue Brain
visualization infrastructure. They are the evolution of existing visualization
systems. Compared to the current single-user or single-presenter systems,
collaborative display systems enable real team work through a combination of
size, resolution and user friendly implementation. Compared to immersive
visualization systems like the CAVE, they provide a simple to use environment
for high-fidelity visualization. For all use cases, the increased display size
and resolution allows improved data exploration for 2D and 3D content.

@@

The core of the Blue Brain visualization infrastructure are multiple
high-resolution tiled display walls driven by our Tide software~\cite{tide}. All
walls are equipped with a multitouch user interface and can be remote controlled
from any web browser. The walls are build using thin-bezel, 55 inch, Full HD LCD
panels with a hardened glass sheet. We use $4\times 3$ and $3\times 3$
configurations for a total of 24 and 18 Megapixel resolution, respectively. The
display size of over five meter diagonally (four meter for $3\times 3$) allows
team-size collaboration (up to ten people) or project-wide presentations (up to
a hundred people). \fig{fTDW} shows one wall during a project-wide presentation
with multiple interactive applications.

@@

In the last years we created a production-ready setup of monoscopic tiled
display walls. On one hand, while there is substantial research on tiled display
wall software, we found that most solutions where not ready to be used in
$24\times 7$ unattained environment. On the other hand, the technology has been
commoditized to make these type of installations affordable to medium-sized
institutions which allowed us to build the software integration for a reasonable
startup cost. The multitouch user interface implements a low entry barrier for
new users, which is a unique capability of our solution.

@@

OpenDeck is our next-generation visualization system, aiming to integrate the
success of tiled display walls with a seamless transition to fully immersive
environments. We are currently in the process of installing the system which
consists of a semi-cylindrical back-projection screen with a 41 Megapixel usable
resolution on a 36 $m^2$ surface (\fig{fOpenDeck}). Like the display walls, it
is equipped with multitouch capabilities which makes it usable as a monoscopic
collaboration system from the first day of installations. Unlike tiled display
walls, it is active stereo capable and is equipped with a 3D tracking system for
immersive rendering. For increased immersion, a lower resolution front
projection system fills in the floor area. OpenDeck will allow us to run our
applications based on Equalizer~\cite{EMP:09} once the system is installed.

@@

The OpenDeck will provide a unique environment for the research and development
of new visualization techniques. The main area of research will be mixed reality
systems, such as touch interface implementation in an immersive environment, and
the transition from 2D content to full immersion. The OpenDeck immersive
infrastructure with multitouch will open a set of questions along immersive
touch user interfaces, transitions and mixing of monoscopic to immersive usage,
the combination of tracked and touch devices, multi-user immersion, latency for
remote immersive rendering as well as multi-site collaboration.

@@

All Blue Brain applications integrate a messaging framework which allows them
to be used as services in a variety of use cases. For example, the Tide web
server providing the user interface shown in \fig{fTideWeb}, is based on this
messaging solution. Other use cases are remote python APIs, JavaScript user
interfaces and service architectures combining multiple visualization
applications with data providers such as HPC simulations.

@@

\begin{figure*}[ht]\center
  \includegraphics[width=\columnwidth]{images/ZeroMS}
  \caption{\label{fUML}UML Diagram of the main messaging classes}
\end{figure*}

@@

The publish-subscribe service is implemented in a \textsf{Publisher} and
\textsf{Subscriber} class. It provides event-based messaging, based on a 128-bit
message type with arbitrary payload. The message type is used for message
subscription, filtering and routing. The payload is expected to be uniquely
identified by the message type, that is, all applications agree for the decoding
and semantics of any given message type. The ZeroBuf section outlines how this
is implemented. The underlying transport uses ZeroMQ pub-sub sockets.

@@

The http server is introspectable, it allows querying the available endpoints
(objects) and the JSON schema~\cite{jsonschema} for each endpoint.
\lref{lhttpReg} shows an excerpt of the Tide registry, and \lref{lhttpSchema} an
excerpt of the schema for one of the exposed objects. This REST API is used by
the Tide web interface from Javascript and to generate remote python APIs.

@@

@@

>>>>

\usepackage{cite}
\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{times}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{caption}
\usepackage{multirow}
\usepackage{paralist}
\usepackage{listings}
\usepackage{wrapfig}

@@

\setlength{\columnsep}{2em}% wrapfig margin, among others

@@

\section{High-Fidelity Displays}

High-fidelity display systems are the integration point of the Blue Brain
visualization infrastructure. They are the evolution of existing visualization
systems, enabling higher resolution, immersion and collaboration. Compared to
the current single-user or single-presenter systems, collaborative display
systems allow real team work through a combination of size, resolution and user
friendly implementation. Compared to immersive visualization systems like the
CAVE, they provide a simple to use environment for high-fidelity visualization.
For all use cases, the increased display size and resolution allows improved
data exploration for 2D and 3D content.

@@

The core of the Blue Brain visualization infrastructure are multiple
high-resolution tiled display walls driven by our Tide software~\cite{tide}. All
walls are equipped with a multitouch user interface and can be remote controlled
from any web browser. The walls are build using thin-bezel, 55 inch, Full HD LCD
panels with a hardened glass sheet. We use $4\times 3$ and $3\times 3$
configurations for a total of 24 and 18 Megapixel resolution, respectively. The
display size of over five meter diagonally (four meter for $3\times 3$) allows
team-size collaboration (up to ten people) or project-wide presentations (up to
a hundred people). \fig{fTDW} shows one wall during a project-wide presentation
with multiple interactive applications running concurrently on the wall.

@@

In the last years we created a production-ready setup of monoscopic tiled
display walls. On one hand, while there is substantial research on tiled display
wall software, we found that most solutions where not ready to be used in
$24\times 7$ unattained environment. For this reason, we started with the TACC
DisplayCluster open source software~\cite{DisplayCluster}, which we
incrementally refactored and improved to the current TIDE implementation.

On the other hand, the technology has been commoditized to make these type of
installations affordable to medium-sized institutions which allowed us to build
the software integration for a reasonable startup cost. We have focused on the
multitouch user interface, which implements a low entry barrier for new users,
a unique capability of our solution.

@@

OpenDeck is our next-generation visualization system, aiming to integrate the
success of tiled display walls with a seamless transition to fully immersive
environments. We are currently in the process of installing the system which
consists of a semi-cylindrical back-projection screen with 41 Megapixel usable
resolution on a 36 $m^2$ surface (\fig{fOpenDeck}). Like the display walls, it
is equipped with multitouch capabilities which makes it usable as a monoscopic
collaboration system from the first day of installation. Unlike tiled display
walls, it is active stereo capable and equipped with a 3D tracking system for
immersive rendering. For increased immersion, a lower resolution front
projection system fills in the floor area. OpenDeck will allow us to run our
applications based on Equalizer~\cite{EMP:09} once the system is installed.

@@

The OpenDeck will provide a unique environment for the research and development
of new visualization techniques. The OpenDeck immersive infrastructure with
multitouch will open a set of questions along immersive touch user interfaces,
transitions and mixing of monoscopic to immersive usage, the combination of
tracked and touch devices, multi-user immersion, latency for remote immersive
rendering as well as multi-site collaboration.

@@

\begin{wrapfigure}{r}{.382\textwidth}\center\vspace{-6ex}
  \includegraphics[width=.382\textwidth]{images/ZeroMS}
  {\caption{\label{fZero}Messaging libraries and use cases}}\vspace{-2ex}
\end{wrapfigure}

All Blue Brain applications integrate a messaging libraries which allow them
to be used as services in a variety of use cases. For example, the Tide web
server providing the user interface shown in \fig{fTideWeb}, is based on this
messaging solution. Other use cases are remote python APIs, JavaScript user
interfaces and service architectures combining multiple visualization
applications with data providers such as HPC simulations (\fig{fZero}).

@@

\begin{figure}[ht]\center
  \includegraphics[width=\columnwidth]{images/ZeroMSUML}
  \caption{\label{fUML}UML Diagram of the main messaging classes}
\end{figure}

@@

The publish-subscribe service is implemented in a \textsf{Publisher} and
\textsf{Subscriber} class. It provides event-based messaging, based on a 128-bit
message type with arbitrary payload. The message type is used for message
subscription, filtering and routing. The payload is expected to be uniquely
identified by the message type, that is, all applications agree for the decoding
and semantics of any given message type. ZeroBuf provides a sample
implementation for this. The underlying transport uses ZeroMQ pub-sub sockets.

@@

The data served by the http server is introspectable, it allows querying the
available endpoints (objects) and the JSON schema~\cite{jsonschema} for each
endpoint. \lref{lhttpReg} shows an excerpt of the Tide registry, and
\lref{lhttpSchema} an excerpt of the schema for one of the exposed objects. This
REST API is used by the Tide web interface from Javascript and to generate
remote python APIs.

@@

\subsection{ZeroBuf}

ZeroBuf is a sample implementation of serialization for ZeroEQ. Based on a
grammar closely related to Flatbuffers schemas~\cite{flatbuffers}, it generates
C++ classes with random \textsf{set/get} access. All data is stored internally
in one continuous memory buffer, which can be used for the ZeroEQ binary
serialization without any copy. The conversion to and from a JSON representation
involves a copy using a \textsf{std::string}. ZeroBuf can store:

\begin{compactitem}
\item (u)int[8,16,32,64,128]\_t, float, double and string members
\item fixed size arrays and dynamic vectors of static-sized elements (intrinsic members or composite types)
\item static and dynamic sub-classes (composite types of the above)
\end{compactitem}

\begin{figure}[ht]\center
  \includegraphics[width=\columnwidth]{images/ZeroBufStatic}\\\vspace{4ex}
  \includegraphics[width=\columnwidth]{images/ZeroBufDynamic}
  \caption{\label{fZeruBuf}ZeroBuf Examples for static (top) and dynamic (bottom) sized objects}
\end{figure}

@@

\fig{fJupyter} shows an example session of using this Python API from a Jupyter
notebook, allocating and launching a Brayns instance, setting relevant
rendering parameters and retrieving an image.

\begin{figure}[ht]\center
  \includegraphics[width=\columnwidth]{images/jupyter}
  \caption{\label{fJupyter}Example Jupyter notebook session using Brayns}
\end{figure}
